---
title: "FTDR Assignment 1: Logistic Regression"
author: "Kladbestand van Jesse Nieuwkoop (1689959)"
output:
  html_document:
    toc: true
    toc_float: true
    df_print: paged
---

---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Course libraries
library(ggplot2)
library(tidyverse)
library(magrittr)
library(micemd)
library(jomo)
library(pan)
library(lme4)
library(knitr)
library(rmarkdown)
library(plotly)
library(devtools)
library(class)
library(car)
library(MASS)
library(ISLR)
library(mice)

# Added (Jesse)
library(car)         # For  IVF test
library(ggcorrplot)  # For correlation matrix heatmap plot
library(patchwork)   # For combining plots
```

# Data preprocessing

## Read dataset

```{r}
# Load dataset
data_raw <- read.csv("../Airbnb_Open_Data.csv")

# Set seed
set.seed(42)

# View dimensions of dataset
dim(data_raw)
```

## Inspect dataset

```{r}
# View first rows of dataset
head(data_raw)
```
```{r}
# View dataset structure
dplyr::glimpse(data_raw)
```

## Filter dataset

```{r}
data_filtered <- data_raw %>%
  dplyr::select(cancellation_policy,
                room.type,
                Construction.year,
                service.fee,
                minimum.nights,
                number.of.reviews,
                reviews.per.month,
                review.rate.number,
                availability.365,
                price)
```

## Rename variables

```{r}
# Rename column names to snake_case format
data_renamed <- data_filtered %>%
  dplyr::rename(
    room_type = room.type,
    construction_year = Construction.year,
    service_fee = service.fee,
    minimum_nights = minimum.nights,
    number_of_reviews = number.of.reviews,
    reviews_per_month = reviews.per.month,
    review_rate_number = review.rate.number,
    availability_365 = availability.365,
    price_usd = price
  )
```

## Missing data handling

```{r}
# Count missing data
data_renamed %>%
  dplyr::summarise(across(everything(), ~ sum(is.na(.)))) %>%
  tidyr::pivot_longer(
    cols = everything(),
    names_to = "variable",
    values_to = "na_count"
  )
```

```{r}
# Determine missing data
total_na_before <- sum(is.na(data_renamed))
nrow_before <- nrow(data_renamed)

# Drop all rows that contain empty or missing values
data_cleaned <- data_renamed %>%
  # Convert empty strings to NA
  mutate(across(where(is.character), ~ trimws(.))) %>%
  mutate(across(where(is.character), ~ na_if(., ""))) %>%
  # Delete all records containing NA values
  drop_na()

# Determine dropped records
nrow_after <- nrow(data_cleaned)
nrow_diff <- nrow_before - nrow_after
total_na_after <- sum(is.na(data_cleaned))

# Inspect result
cat(total_na_before, "missing values\n")
cat(nrow_diff, "dropped rows\n")
cat(total_na_after, "resulting missing values\n")
```

## Data type convertion

```{r}
# Determine factor levels
table(data_cleaned$cancellation_policy)
table(data_cleaned$room_type)
```

```{r}
# Convert necessary variables to factor types
data_converted <- data_cleaned %>%
  dplyr::mutate(
    cancellation_policy = factor(cancellation_policy, 
                                 levels = c("flexible", "moderate", "strict"),
                                 ordered = TRUE),
    room_type = factor(room_type, 
                       ordered = FALSE)
  )
```

```{r}
# Verify structure of factors
str(data_converted$cancellation_policy)
str(data_converted$room_type)

# Verify levels of factors
levels(data_converted$cancellation_policy)
levels(data_converted$room_type)

# Verify if ordered
is.ordered(data_converted$cancellation_policy)
is.ordered(data_converted$room_type)
```

## Data parsing

```{r}
# Parse the price and service fee variables
data_parsed <- data_converted %>%
  dplyr::mutate(
    price_usd = parse_number(price_usd),
    service_fee = parse_number(service_fee)
  )
```

```{r}
# Inspect updated data structure
str(data_parsed$price_usd)
str(data_parsed$service_fee)
```

## Outlier handling

```{r}
data_pre_outlier_rmvl <- data_parsed

numeric_columns <- c("construction_year", 
                     "service_fee", 
                     "minimum_nights", 
                     "number_of_reviews", 
                     "reviews_per_month", 
                     "review_rate_number", 
                     "availability_365", 
                     "price_usd")
```

### General

#### Visualization summarization function

```{r}
plot_info <- function(data, column_name) {
  # Ensure the input column exists in the dataset
  if (!column_name %in% names(data)) {
    stop(paste("Column", column_name, "not found in the dataset."))
  }
  
  # Extract the column
  column_data <- data[[column_name]]
  
  # Ensure the column is numeric
  if (!is.numeric(column_data)) {
    stop(paste("Column", column_name, "is not numeric."))
  }
  
  # Histogram
  hist_plot <- ggplot(data, aes(x = !!sym(column_name))) +
    geom_histogram(bins = 30, fill = "skyblue", color = "black") +
    labs(title = paste("Histogram of", column_name), x = column_name, y = "Frequency") +
    theme_minimal()
  
  # Boxplot
  box_plot <- ggplot(data, aes(x = !!sym(column_name))) +
    geom_boxplot(fill = "skyblue", color = "black") +
    labs(title = paste("Boxplot of", column_name), x = column_name, y = "") +
    theme_minimal()
  
  # Density Plot
  density_plot <- ggplot(data, aes(x = !!sym(column_name))) +
    geom_density(fill = "skyblue", alpha = 0.5) +
    labs(title = paste("Density Plot of", column_name), x = column_name, y = "Density") +
    theme_minimal()
  
  # Q-Q Plot
  qq_plot <- ggplot(data, aes(sample = !!sym(column_name))) +
    stat_qq(color = "skyblue") +
    stat_qq_line(color = "red") +
    labs(title = paste("Q-Q Plot of", column_name), x = "Theoretical Quantiles", y = "Sample Quantiles") +
    theme_minimal()
  
  # Combine all plots into a 2x2 grid
  combined_plot <- (hist_plot | density_plot) / (box_plot | qq_plot)
  
  # Print the combined plot
  print(combined_plot)
}

```

#### Negative numeric removal

```{r}
# Count total rows
nrow_bef_negremoval <- nrow(data_pre_outlier_rmvl)

# Remove all rows that contain values lower than 0
data_outlier_0 <- data_pre_outlier_rmvl %>%
  dplyr::filter(
    service_fee >= 0,
    minimum_nights >= 0,
    number_of_reviews >= 0,
    reviews_per_month >= 0,
    review_rate_number >= 0,
    availability_365 >= 0,
    price_usd >= 0
  )

# Print n removed rows
cat(nrow_bef_negremoval - nrow(data_outlier_0), "rows removed")
```

#### IQR

```{r}
detect_outliers_iqr <- function(data, columns) {
  for (col in columns) {
    # Calculate Q1, Q3, and IQR
    Q1 <- quantile(data[[col]], 0.25, na.rm = TRUE)
    Q3 <- quantile(data[[col]], 0.75, na.rm = TRUE)
    IQR <- Q3 - Q1
    
    # Calculate lower and upper bounds
    lower_bound <- Q1 - 1.5 * IQR
    upper_bound <- Q3 + 1.5 * IQR
    
    # Print bounds for the variable
    cat("---", col, "---\n")
    cat("Q1:", Q1, "Q3:", Q3, "IQR:", IQR, "\n")
    cat("Lower Bound:", lower_bound, "Upper Bound:", upper_bound, "\n")
    
    # Count outliers
    num_outliers <- sum(data[[col]] < lower_bound | data[[col]] > upper_bound, na.rm = TRUE)
    cat("Number of outliers:", num_outliers, "\n\n")
    
    # Filter out outliers
    data <- data[data[[col]] >= lower_bound & data[[col]] <= upper_bound, ]
  }
  return(data)
}
```

```{r}
# Apply IQR outlier removal
data_iqr <- detect_outliers_iqr(data_outlier_0, numeric_columns)
```

### Specific

#### 1: construction_year

Ziet er prima uit: Geen extreme waarden / outliers.

```{r}
# Print summary
summary(data_outlier_0$construction_year)
```
```{r}
# Plot summary info
plot_info(data_outlier_0, "construction_year")
```

```{r}
data_outlier_1 <- data_outlier_0
```

#### 2: (service_fee)

Ziet er prima uit: Geen extreme waarden / outliers.

```{r}
# Print summary
summary(data_outlier_1$service_fee)
```
```{r}
# Plot summary info
plot_info(data_outlier_1, "service_fee")
```

```{r}
data_outlier_2 <- data_outlier_1
```

#### 3: minimum_nights

Veel extreme waarden: Komt waarschijnlijk omdat de hoge waarden ingesteld worden, zodat mensen niet gaan boeken (Zie bronnen). Geen idee hoe hier mee om te gaan...

1. Boven 90 dagen alles weghalen (maar hier haal je ook legitieme values mee weg)
2. Boven 365 dagen alles weghalen (maar hiermee weinig verschil)

[Reddit 1](https://www.reddit.com/r/AirBnB/comments/1bg4oic/extremely_long_minimum_stay_policies_upwards_of/)

> "I think it might be there way of blocking those dates while still being available on Airbnb search, I'm not sure. I've also seen the other method where listings that are usually $50 a night be priced at $10,000 a night. I, too, would like a clear explanation."

[Reddit 2](https://www.reddit.com/r/AirBnB/comments/tcmlnw/90_day_minimum_stay_on_airbnb/)

> "I often bump up my minimum night stay to actually prevent bookings during a certain time frame. Then one day, to my absolute shock, someone actually booked. So maybe just send them a message, it could be way too many different things."

> "That’s local government limit. Some government requires the listing to host short term stays only with a permit. If the host didn’t get the permit in time, the listing is enforced to long term stays only. The threshold is per the regulation, generally 30, 90, or even 365 days. Besides, France government also permits a special kind of short term rental, mobility lease, which is 30 days minimum without the permit."

```{r}
# Print summary
summary(data_outlier_2$minimum_nights)
```
```{r}
# Plot summary info
plot_info(data_outlier_2, "minimum_nights")
```

```{r}
# Print total amount of > 370 minimum_nights
data_outlier_2 %>%
  dplyr::filter(minimum_nights > 370) %>%
  dplyr::summarise(count = n())
```
```{r}
# Remove extreme values with upper threshold
data_outlier_TEMP2 <- data_outlier_2 %>%
  dplyr::filter(minimum_nights <= 95)
```

```{r}
# Print summary
summary(data_outlier_TEMP2$minimum_nights)
```
```{r}
# Plot summary info
plot_info(data_outlier_TEMP2, "minimum_nights")
```
```{r}
# Apply IQR outlier removal
data_iqr <- detect_outliers_iqr(data_outlier_TEMP2, "minimum_nights")
```

```{r}
data_outlier_3 <- data_outlier_2
```

#### 4: number_of_reviews

Deze verdeling is vrij logisch. Ik zie geen reden tot outlier removal. Ook months_active laat geen verdachte verdeling zien.

```{r}
# Print summary
summary(data_outlier_3$number_of_reviews)
```
```{r}
# Plot summary info
plot_info(data_outlier_3, "number_of_reviews")
```
```{r}
# Print total amount of reviews > 100
data_outlier_3 %>%
  dplyr::filter(number_of_reviews > 100) %>%
  dplyr::summarise(count = n())
```

```{r}
# Create extra column that divides number of reviews by reviews per month
data_outlier_3 <- data_outlier_3 %>%
  dplyr::mutate(months_active = number_of_reviews / reviews_per_month)
```

```{r}
# Print summary
summary(data_outlier_3$months_active)
```
```{r}
# Plot summary info
plot_info(data_outlier_3, "months_active")
```

```{r}
data_outlier_4 <- data_outlier_3
```

#### 5: reviews_per_month

Door andere info in the grafiek te gebruiken, zijn er veel datapunten die in een maand meer reviews per maand ontvangen dan mogelijk n.a.v. max_stays_per_month uit te rekenen. Probleem hiermee is natuurlijk dat het mogelijk is dat een stay bijvoorbeeld in een latere maand gereviewed kan worden. Maar dan nog, als een stay 31 boekingen per maand kan hebben, betekent dit dat reviews_per_month

```{r}
# Print summary
summary(data_outlier_4$reviews_per_month)
```
```{r}
# Plot summary info
plot_info(data_outlier_4, "reviews_per_month")
```

```{r}
# Create new column 'max_stays_per_month'
data_outlier_4 <- data_outlier_4 %>%
  dplyr::mutate(max_stays_per_month = 31 / minimum_nights)
```

```{r}
# Print summary
summary(data_outlier_4$max_stays_per_month)
```
```{r}
# Plot summary info
plot_info(data_outlier_4, "max_stays_per_month")
```
```{r}
# Count the points above the line y = x
count_above_line <- sum(data_outlier_4$reviews_per_month > data_outlier_4$max_stays_per_month, na.rm = TRUE)

# Scatterplot with y = x line, colored points, and count in title
ggplot(data_outlier_4, aes(x = max_stays_per_month, y = reviews_per_month)) +
  geom_point(aes(color = reviews_per_month > max_stays_per_month), alpha = 0.6) +  # Color points based on condition
  geom_abline(slope = 1, intercept = 0, color = "black", linetype = "dashed", size = 1) +  # y = x line
  scale_color_manual(values = c("FALSE" = "blue", "TRUE" = "red"), labels = c("Below Line", "Above Line")) +  # Define colors
  labs(
    title = paste("Reviews per Month vs. Maximum Stays per Month\nPoints Above Line:", count_above_line),
    x = "Maximum Stays per Month",
    y = "Reviews per Month",
    color = "Point Position"
  ) +
  theme_minimal()
```

```{r}
# Remove data_outlier_pos$reviews_per_month > data_outlier_pos$max_stays_per_month
data_outlier_5 <- data_outlier_4 %>%
  dplyr::filter(reviews_per_month <= max_stays_per_month)
```

#### 6: review_rate_number

Niets sus. Geen uitschieters.

```{r}
# Print summary
summary(data_outlier_5$review_rate_number)
```
```{r}
# Plot summary info
plot_info(data_outlier_5, "review_rate_number")
```

```{r}
data_outlier_6 <- data_outlier_5
```

#### 7: availability_365

```{r}
# Print summary
summary(data_outlier_6$availability_365)
```
```{r}
# Plot summary info
plot_info(data_outlier_6, "availability_365")
```

```{r}
data_outlier_6 %>%
  dplyr::arrange(desc(availability_365)) %>%  # Sort rows by availability_365 in descending order
  dplyr::select(availability_365)            # Select only the availability_365 column
```

```{r}
# remove all values above 1000 in availability_365
data_outlier_7 <- data_outlier_6 %>%
  dplyr::filter(availability_365 <= 1000)
```

#### 8: price_usd

In principe geen uitschieters, maar wel opvallend dat de data niet normaal verdeeld is rond "normale" bedragen, alhoewel... verblijf is vaak duur.

```{r}
# Print summary
summary(data_outlier_7$price_usd)
```
```{r}
# Plot summary info
plot_info(data_outlier_7, "price_usd")
```

```{r}
# Plot summary info
data_outlier_8 <- data_outlier_7
```

## Final inspection

```{r}
# Store last determined dataset in separate object
data_inspect <- data_outlier_8
```

```{r}
# Count missing data
data_inspect %>%
  dplyr::summarise(across(everything(), ~ sum(is.na(.)))) %>%
  tidyr::pivot_longer(
    cols = everything(),
    names_to = "variable",
    values_to = "na_count"
  )
```

```{r}
# Inspect data structure
dplyr::glimpse(data_inspect)
```

```{r}
data_ready <- data_inspect
```

## Save 1000 size sample

```{r}
# Take a random sample of 1000 rows
Airbnb_Open_Data_sample <- data_ready %>%
  sample_n(1000)

# Save the sampled dataset to a CSV file
write.csv(Airbnb_Open_Data_sample, "Airbnb_Open_Data_1000.csv", row.names = FALSE)
```

# Linear Regression

## Prepare data for analysis

```{r}
# Standardize relevant numeric variables
data_standardized <- data_ready %>%
  mutate(
    service_fee = as.numeric(scale(service_fee)),
    minimum_nights = as.numeric(scale(minimum_nights)),
    number_of_reviews = as.numeric(scale(number_of_reviews)),
    reviews_per_month = as.numeric(scale(reviews_per_month)),
    review_rate_number = as.numeric(scale(review_rate_number)),
    availability_365 = as.numeric(scale(availability_365))
  )
```

```{r}
# Check mean and standard deviation for scaled variables
summary(data_standardized)
```
```{r}
summary(data_standardized$minimum_nights)
summary(data_standardized$availability_365)

```
```{r}
library(ggplot2)

# Plot for minimum_nights
ggplot(data_ready, aes(x = minimum_nights)) +
  geom_histogram(bins = 50, fill = "blue", alpha = 0.7) +
  labs(title = "Distribution of Minimum Nights", x = "Minimum Nights", y = "Count")

# Plot for availability_365
ggplot(data_ready, aes(x = availability_365)) +
  geom_histogram(bins = 50, fill = "green", alpha = 0.7) +
  labs(title = "Distribution of Availability (365)", x = "Availability (365)", y = "Count")

```

```{r}
# Quantiles for extreme values
quantile(data_standardized$minimum_nights, probs = c(0.95, 0.99))
quantile(data_standardized$availability_365, probs = c(0.95, 0.99))

```

```{r}
# Set active data to last processed set
data <- data_standardized
```

```{r}
library(dplyr)

data_ready %>%
  filter(minimum_nights > 100) %>%
  summarise(count = n())

```


```{r}
dplyr::glimpse(data)
```

## Prepare functions

```{r}
test_multicol <- function(model) {
  # Ensure the model is valid
  if (!inherits(model, "lm")) {
    stop("Input must be a linear model (lm object).")
  }
  
  # Step 1: Show VIF results
  print(car::vif(model))
  
  # Step 2: Extract data from the model
  model_data <- model$model
  
  # Ensure only numeric columns are used for the correlation matrix
  numeric_data <- model_data %>%
    select_if(is.numeric)
  
  # Step 3: Show correlation matrix
  cor_matrix <- cor(numeric_data)
  
  # Step 4: Visualize correlation matrix
  print(ggcorrplot::ggcorrplot(cor_matrix, 
                         method = "circle", 
                         type = "lower", 
                         lab = TRUE, 
                         title = "Correlation Matrix of Numeric Predictors")
  )
}
```

## Models

### Model 1: price_usd (All)

```{r}
model_1 <- lm(price_usd ~ 
              cancellation_policy + 
              room_type + 
              construction_year + 
              service_fee +
              minimum_nights +
              number_of_reviews +
              reviews_per_month +
              review_rate_number +
              availability_365, 
            data = data_parsed)
summary(model_1)
```
```{r}
# Test multicollinearity
test_multicol(model_1)
```

### Model 2: price_usd (Ex. service_fee, reviews_per_month)

```{r}
model_2 <- lm(price_usd ~ 
              cancellation_policy + 
              room_type + 
              construction_year + 
              minimum_nights +
              number_of_reviews +
              review_rate_number +
              availability_365, 
            data = data_parsed)
summary(model_2)
```
```{r}
# Test multicollinearity
test_multicol(model_2)
```

### Model 3: review_rate_number (Ex. service_fee, reviews_per_month)

```{r}
model_3 <- lm(review_rate_number ~ 
              cancellation_policy + 
              room_type + 
              construction_year + 
              minimum_nights +
              number_of_reviews +
              price_usd +
              availability_365, 
            data = data_parsed)
summary(model_3)
```

```{r}
# Test multicollinearity
test_multicol(model_3)
```
